Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-13 arxiv/diff_cov.tex
typos

The previous section assumes that all tasks have the same sample size and feature vectors.
This section studies how different sample sizes and different features impact hard parameter sharing.
%The different covariates case differs from the same covariates case in two aspects.
%First, different tasks may have different sample sizes. In extreme scenarios, one task may have much less labeled data compared to another task.
The setting where features differ across tasks is often also called ``covariate shift''.

Unlike the previous section, we no longer have an optimal solution for $f(A, B)$.
This is because $f(A, B)$ is in general non-convex.
Instead, our result implies tight generalization bounds for any \emph{local minimum} of $f(A, B)$.
We focus on the two-task case to better understand the impact of having different sample sizes and different covariates.
Suppose that $X^{(1)} = Z^{(1)}(\Sigma^{(1)})^{1/2} \in \real^{n_1 \times p}$ and $X^{(2)} = Z^{(2)}(\Sigma^{(2)})^{1/2} \in \real^{n_2 \times p}$, where the entries of $Z^{(1)}$ and $ Z^{(2)}$ are drawn independently from a one dimensional distribution with zero mean, unit variance, and constant $\varphi$-th moment for a fixed $\varphi > 4$. The matrices $\Sigma^{(1)}\in \R^{p\times p}$ and $\Sigma^{(2)}\in \R^{p\times p}$ denote the population covariance matrices of task 1 and task 2, respectively.

\paragraph{Bias-variance equations.}
Our key result characterizes the asymptotic limit of the inverse of the sum of two different sample covariance matrices.
To motivate our study, consider a special case where $A_1=A_2=1$. 
Without loss of generality, we consider task two's prediction loss and the same result applies to task one.
By Proposition 1 of \citet{WZR20}, we consider the case of $r = 1 < t = 2$, since when $r > 1$, the optimal solution of $f(A, B)$ is equivalent to single-task learning.
When $r = 1$, $B$ is a vector and $A_1, A_2$ are both scalars.
Hence we can write down a closed form equation for any local minimum of $f(A, B)$.
By solving $B$ in equation \eqref{eq_mtl}, we obtain the hard parameter sharing estimator $\hat{\beta}_2^{\MTL} = B A_2 = B$ as follows:
\begin{align}
	\hat{\beta}_2^{\MTL} &= {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \ \ \text{ where } \nonumber \\
	\hat{\Sigma} &= {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}. \label{def hatsig}
\end{align}
The matrix $\hat{\Sigma}$ adds up both tasks' sample covariances, and the expectation of $\hat{\Sigma}$ is equal to a mixture of their population covariance matrices, with mixing proportions determined by their sample sizes.

@@

\begin{theorem}\label{thm_main_RMT}
	Let $c_{\varphi}$ be a fixed value in $(0, \frac{\varphi - 4}{2\varphi})$.
	Assume that: a) the sample sizes $n_1$ and $n_2$ both satisfy Assumption \ref{assume_rm};
	b) $M$'s singular values are all greater than $\tau$ and less than $1/\tau$;
	c) task one's sample size is greater than $\tau p$ and task two's sample size is greater than $(1 + \tau) p$.
	With high probability over the randomness of $X^{(1)}$ and $X^{(2)}$, we have that
	the variance equation \eqref{eq_variance_2task} satisfies the following estimate with high probability:
			\begin{align}\label{lem_cov_shift_eq}
				&\bigabs{\bigtr{\Sigma^{(2)} \bigbrace{\frac{n_1 + n_2}p{\hat{\Sigma}^{-1}} - \frac{(a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}{p} }}} \nonumber \\
			&\le  p^{-c_{\varphi}}, %\norm{\Sigma^{(2)}},
			\end{align}
			where $a_1$ and $a_2$ are the solutions of the following self-consistent equations
			\begin{align}
				&a_1 + a_2 = 1- \frac{p}{n_1 + n_2},\\
				&a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{n_1}{n_1 + n_2}. \label{eq_a12extra}
			\end{align}
\end{theorem}
The bias limit can be found in Section \ref{appendix RMT}.
Our result extends Fact \ref{fact_tr} to the inverse of the sum of two sample covariance matrices.
To see this, when $n_1$ is zero, we solve equation \eqref{eq_a12extra} and obtain $a_1 = 0$ and $a_2 = (n_2-p) / n_2$, and apply these solutions to equation \eqref{lem_cov_shift_eq}.

>>>>

The previous section assumes that all tasks have the same sample size and feature vectors.
In this section, we study how different sample sizes and different features impact hard parameter sharing.
%The different covariates case differs from the same covariates case in two aspects.
%First, different tasks may have different sample sizes. In extreme scenarios, one task may have much less labeled data compared to another task.
The setting where features differ across tasks is often called ``covariate shift''.

Unlike the previous section, we no longer have an optimal solution to $f(A, B)$.
This is because $f(A, B)$ is in general non-convex.
Instead, our result implies tight generalization bounds for any \emph{local minimizer} of $f(A, B)$.
We focus on the two-task case to better understand the impact of having different sample sizes and different covariates.
Suppose that $X^{(1)} = Z^{(1)}(\Sigma^{(1)})^{1/2} \in \real^{n_1 \times p}$ and $X^{(2)} = Z^{(2)}(\Sigma^{(2)})^{1/2} \in \real^{n_2 \times p}$, where the entries of $Z^{(1)}$ and $ Z^{(2)}$ are drawn independently from a one dimensional distribution with zero mean, unit variance, and constant $\varphi$-th moment for a fixed $\varphi > 4$. The matrices $\Sigma^{(1)}\in \R^{p\times p}$ and $\Sigma^{(2)}\in \R^{p\times p}$ denote the population covariance matrices of task 1 and task 2, respectively.

\paragraph{Bias-variance equations.}
Our key result characterizes the asymptotic limit of the inverse of the sum of two different sample covariance matrices.
Without loss of generality, we consider task two's prediction loss and the same result applies to task one.
By Proposition 1 of \citet{WZR20}, we consider the case of $r = 1 < t = 2$, since when $r > 1$, the optimal solution of $f(A, B)$ is equivalent to single-task learning.
When $r = 1$, $B$ is a vector and $A_1, A_2$ are both scalars.
To motivate our study, we consider a special case where $A_1=A_2=1$. 
%Hence we can write down a closed form equation for any local minimizer of $f(A, B)$.
By solving $B$ in equation \eqref{eq_mtl}, we obtain the hard parameter sharing estimator $\hat{\beta}_2^{\MTL} = B A_2 = B$ as follows:
\begin{align}
	\hat{\beta}_2^{\MTL} &= {\hat{\Sigma}}^{-1} ({X^{(1)}}^{\top} Y^{(1)} + {X^{(2)}}^{\top} Y^{(2)}), \ \ \text{ where } \nonumber \\
	\hat{\Sigma} &= {X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}. \label{def hatsig}
\end{align}
The matrix $\hat{\Sigma}$ adds up both tasks' sample covariance matrices, and the expectation of $\hat{\Sigma}$ is equal to a mixture of their population covariance matrices, with mixing proportions determined by their sample sizes.

@@

\begin{theorem}\label{thm_main_RMT}
	Let $c_{\varphi}$ be a fixed value in $(0, \frac{\varphi - 4}{2\varphi})$.
	Assume that: a) the sample sizes $n_1$ and $n_2$ both satisfy Assumption \ref{assume_rm};
	b) $M$'s singular values are all greater than $\tau$ and less than $1/\tau$;
	c) task one's sample size is greater than $\tau p$ and task two's sample size is greater than $(1 + \tau) p$.
	With high probability over the randomness of $X^{(1)}$ and $X^{(2)}$, we have that
	the variance equation \eqref{eq_variance_2task} satisfies the following estimate with high probability:
			\begin{align}\label{lem_cov_shift_eq}
				&\bigabs{\bigtr{\Sigma^{(2)} \bigbrace{\frac{n_1 + n_2}p{\hat{\Sigma}^{-1}} - \frac{(a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}{p} }}} \nonumber \\
			&\le  p^{-c_{\varphi}}, %\norm{\Sigma^{(2)}},
			\end{align}
			where $a_1$ and $a_2$ are the solutions of the following self-consistent equations
			\begin{align}
				&a_1 + a_2 = 1- \frac{p}{n_1 + n_2},\\
				&a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{n_1}{n_1 + n_2}. \label{eq_a12extra}
			\end{align}
\end{theorem}
The bias limit can be found in Section \ref{appendix RMT}.
Our result extends Fact \ref{fact_tr} to the inverse of the sum of two sample covariance matrices.
To see this, when $n_1$ is zero, we solve equation \eqref{eq_a12extra} and obtain $a_1 = 0$ and $a_2 = (n_2-p) / n_2$, and apply these solutions to equation \eqref{lem_cov_shift_eq}.
We also remark that for general $A_1$ and $A_2$ that are not equal to 1, we need to replace $\hat \Sigma$ with $\frac{A_1^2}{A_2^2}{X^{(1)}}^{\top} X^{(1)} + {X^{(2)}}^{\top} X^{(2)}$.
This amounts to replacing $M$ with $\frac{A_1}{A_2}M$ in Theorem \ref{thm_main_RMT}.
